Testable normative theories

### 2020-05-02 simple.tex
Sketch of what comes next

\subsection{Definition}
A normative theory is a tuple $(\allprops, ¬, \allargs, I, (\ileadstost[i])_{i \in I}, q, (\gleadsto[q(i)])_{q(i) \in q(I)}, \gbeats, +)$,
with ${\ileadstost[i]} \subseteq (\allargs × \allargs) × \Phi$, ${\gleadsto[q(i)]} \subseteq \allargs × \Phi$, ${\gbeats} \subseteq \allargs × \allargs$,
where $\Phi = \allprops × \set{\text{possible}, \text{sure}}$, and $+$ a commutative associative internal binary operation on $\allargs$. The notation $\sum \args = \sum_{\ar \in \args} \ar$ thus designates a well-defined argument.

@@

\subsection{Theorem}
\begin{theorem}
	If a theory is answerable, has bounded length, does not self-attack, and is operationally valid, it is valid.
\end{theorem}
\begin{proof}
	Lemma: given $\phi \in {\gleadsto}(\allargs)$ and $\ar[k] \in \gargs[\phi]$, if $\ar[k] \nileadstost \phi$, then: $\exists \ar[k + 1] \in \gargs[\phi] \suchthat \card{c(\ar[k + 1])} ≥ \card{c(\ar[k])} + 1$.

	Pick any $\phi \in {\gleadsto}(\allargs)$ and $\ar[k] \in \gargs[\phi]$ such that $\ar[k] \nileadstost \phi$. We know that $\exists \ar_1 \in \allargs \suchthat \ar_1 \ibeatse[!\phi] \ar[k]$. Therefore, by Answerability, $\exists \ar_1' \in \allargs \suchthat \ar_1' \gbeatsstar \ar[k] \land \ar_1' \ibeatse[!\phi] \ar[k]$, and, by Operational validity, $\exists \ar_2 \in \gargs[\phi] \suchthat \ar_2 \gbeats \ar'_1$ and $\ar[k] + \ar_2 \ibeatse[\phi] \ar'_1$. 
	Define $\ar[k + 1] = \ar_2 + \ar[k]$.

	\commentOC{Note that we do not have $\ar_1 \gbeats \ar[k]$.}
	By No self-attack, $¬(\ar[k] \gbeatsstar \ar'_1)$. Thus, $\ar_2 \notin c(\ar[k])$ (because $\ar_2 \in c(\ar[k])$ and $\ar_2 \gbeats \ar'_1$ implies $\ar[k] \gbeatsstar \ar'_1$ by definition of $\gbeatsstar$).
	Thus, $\card{c(\ar[k + 1])} ≥ \card{c(\ar[k])} + 1$.

@@

>>>>

\subsection{Definition}
A normative theory is a tuple $(\allprops, ¬, \allargs, I, (\ileadstost[i])_{i \in I}, q, (\gleadsto[q(i)])_{q(i) \in q(I)}, \gbeats, +)$,
with ${\ileadstost[i]} \subseteq (\allargs × \allargs) × \Phi$, ${\gleadsto[q(i)]} \subseteq \allargs × \Phi$, ${\gbeats} \subseteq \allargs × \allargs$,
where $\Phi = \allprops × \set{\text{possible}, \text{sure}}$, and $+$ a commutative associative internal binary operation on $\allargs$. Given $\args \subseteq \allargs$, the notation $\sum \args = \sum_{\ar \in \args} \ar$ thus designates a well-defined argument.

@@

\subsection{Theorem}
\begin{theorem}
	If a theory is answerable, has bounded length, does not repeat, and is operationally valid, it is valid.
\end{theorem}
\begin{proof}
	Lemma: given $\phi \in {\gleadsto}(\allargs)$ and $\ar[k] \in \gargs[\phi]$, if $\ar[k] \nileadstost \phi$, then: $\exists \ar[k + 1] \in \gargs[\phi] \suchthat \card{c(\ar[k + 1])} ≥ \card{c(\ar[k])} + 1$.

	Pick any $\phi \in {\gleadsto}(\allargs)$ and $\ar[k] \in \gargs[\phi]$ such that $\ar[k] \nileadstost \phi$. We know that $\exists \ar_1 \in \allargs \suchthat \ar_1 \ibeatse[!\phi] \ar[k]$. Therefore, by Answerability, $\exists \ar_1' \in \allargs \suchthat \ar_1' \gbeatsstar \ar[k] \land \ar_1' \ibeatse[!\phi] \ar[k]$, and, by Operational validity, $\exists \ar_2 \in \gargs[\phi] \suchthat \ar_2 \gbeats \ar'_1$ and $\ar[k] + \ar_2 \ibeatse[\phi] \ar'_1$. 
	Define $\ar[k + 1] = \ar[k] + \ar_2$.

	\commentOC{Note that we do not have $\ar_1 \gbeats \ar[k]$.}
	By No repetition, $\ar_2 \notin c(\ar[k])$.
	Thus, $\card{c(\ar[k + 1])} ≥ \card{c(\ar[k])} + 1$.

@@

\subsection{Current question}
Should I leave open the uplifting of $+$ for more generality? Is this really sufficiently general to encompass both interpretations of Answerability, and more? Or should I stick to the current, “single-try”, approach and interpretation?

\subsection{Interpreting Answerability}
At a given stage in operationalizing the protocol, define $f(\ar_0)$ as the arguments that are considered as having a chance to attack $\ar_0$ (without backtracking). Thus, this will not include $\ar_1$ if indeed $\ar_2$ is considered by $i$ as a satisfying answer to $\ar_1$. We can interpret Answerability in two ways. Strong answ: $\exists \ar_1 \in f(\ar_0) \suchthat \ar_1 \gbeats \ar_0$. Weak answ: $f(\ar_0) ≠ \emptyset ⇒ \exists \ar_1' \suchthat \ar_1' \gbeats \ar_0$.

A model can be allowed to try only once each answer. If $\ar_2 \gbeats \ar_1 \land \ar_1 \in f(\ar_0 + \ar_2)$, the model has lost.

Thus, four variants, given $\ar_2 \gbeats \ar_1$.
\begin{itemize}
	\item Strong answ with single try: $\ar_1 \notin f(\ar_0 + \ar_2) \land \forall \ar_1' \in f(\ar_0 + \ar_2): \ar_1' \gbeatsstar \ar_0 + \ar_2$.
	\item Strong answ with multiple tries: $\forall \ar_1' \in f(\ar_0 + \ar_2): \ar_1' \gbeatsstar \ar_0 + \ar_2$. (Example: $\ar_1 = \text{“for ethical reasons”}$, and $\eta$ knows two c-a to this, from different angles.) 	\item Weak answ with single try: $\ar_1 \notin f(\ar_0 + \ar_2) \land [f(\ar_0 + \ar_2) ≠ \emptyset ⇒ \exists \ar_1' \suchthat \ar_1' \gbeats \ar_0]$.
	\item Weak answ with multiple tries: $f(\ar_0 + \ar_2) ≠ \emptyset ⇒ \exists \ar_1' \suchthat \ar_1' \gbeats \ar_0 + \ar_2$.
\end{itemize}

Goal: leave addition free so as to admit multiple tries, and prove that Strong answ with multiple tries is the right one for a general approach: suitable restriction to the model permits single try (with same conditions); and weak answ actually does not permit more models.

When $f$ is available and is single-valued, question: should we use Strong or weak answ? Right answer: use Strong answ. It is stricly more general. A model that wants weak answ can define its internal attacks suitable (to be proven); and a model that wants to claim more can.

When $f$ is not available, only remains: single try (just give a decisive argument) or multiple tries (try more and more arguments, blindly).

Suppose that a model also defines $f_\eta$, meaning: $f_\eta(\ar_1) = \ar_2$ the argument the model plays against $\ar_1$. Thus $f_\eta$, like $\gbeats$, is a binary relation over $\allargs$, but $f_\eta$ in supplement returns a single argument or none. Also, $\ar_2 = f(\ar_1) ⇒ \ar_2 \gbeats \ar_1$. When a reason $\ar_1$ for rejecting $\ar_0$ is not known, $f_\eta(\ar_0)$ instead gives the argument to be played next. Note that $f_\eta$ is only useful when (in the known c-a variant) for some $\ar_1$ there are multiple candidates $\ar_2$ or when (in the unknown c-a variant) for some $\ar_0$ here are multiple candidates $\ar_2$. In fact, this is more complicated: $f_\eta(\ar_1)$ can give first $\ar_2$, then, if that is not enough, $\ar_4$ (thus it’s rather $f_\eta(\ar_1, \ar_0)$).

Under Strong anws, we ask that $f(\ar_0) \cap {\gbeatsinv}(\ar_0) ≠ \emptyset$, and (inevitably) that $f_\eta(\ar_1) ≠ \emptyset$. Under weak answ, we do not expect anticipation and only demand that $f_\eta(\ar_1) ≠ \emptyset$. But this does not allow supplementary models. TODO prove this.

\subsection{Single-try models}
The above models can define several $\ar_2$ for one $\ar_1$. Even when interpreting Answerability in the strong sense (mandating that exactly the $\ar_1'$ we search for be defined, not just any $\ar_1'$), this gives a lot of flexibility to the model, and correspondingly, makes it very difficult to falsify: we only know that is unconclusive when all possible subset of answers to a $\ar_1$ has been tried

TODO false: we try any sequence and then poof. Thus, is my approach insufficiently general (and thus should be restrained to the single-try models anyway)?
